Concept note

### 2018-04-28 inst/rmd/externalities.Rmd
big update

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 3,
               fig.width = 4,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h")
opts_knit$set(width=75)
options(xtable.comment = FALSE)

@@

To test whether IRS from one household has an effect on workers living in another household, we first must make some assumptions about the level and waning effect of IRS generally. For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009]. We simplify these assumptions into a time-protection matrix, which can be visualized in the below chart.

```{r}
protection <- 
    data_frame(months_since = c('No IRS / > 1 yr', '01', '02-04', '05-09', '10-12'),
               protection = c(0, 1, 3, 2, 1))
ggplot(data = protection,
       aes(x = months_since,
           y = protection,
           group = 1)) +
  geom_area(fill = 'lightblue',
            alpha = 0.6) +
  geom_line() +
  theme_maragra() +
  labs(x = 'Months since IRS',
       y = '"Protection" level')
```

@@

```{r, echo = TRUE}
weighter <- function(x){
    x[x == 0] <- 0.01
    out <- (1 / x)^1.2
    return(out)
}
```

Where `x` is the distance (in kilometers) to the household whose "herd protection" score is being estimated. In other words, a household's "herd protection" score is the mean of all households' protection scores, weighted by the inverse of the distance away raised to the power of 1.2.

@@

@@

@@

>>>>

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)

@@

To test whether IRS from one household has an effect on workers living in another household, we first must make some assumptions about the level and waning effect of IRS generally. For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [@White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009]. We simplify these assumptions into a time-protection matrix, which can be visualized in the below chart.

```{r}
x <- c(1:6)
y <- (c(-3.45, -6.04, - 6.53, -0.17, -2.74, 5.27) + 27.61) / 27.61
lo <- loess(y~x, span = 3)$fitted
lo[lo > 1] <- 1
lo <- 1-lo
y <- 1-y
df <- data.frame(x,y,lo)
df$months_since <- add_zero(df$x, n = 2)
df$months_since[df$months_since == '06'] <- '06+'
df <- df %>%
  bind_rows(df[nrow(df),] %>%
              mutate(months_since = 'No IRS'))
df <- df %>%
  gather(key, value, y:lo) %>%
  mutate(key = ifelse(key == 'lo', 
                      'Brew', 'Tukei'))
ggplot(data = df,
       aes(x = x,
           y = value,
           group = key,
           color = key)) +
  geom_point() +
  geom_line() +
  # geom_smooth(se = FALSE,span = 3) +
  theme_maragra() +
  labs(x = 'Months after IRS',
       y = 'Protection') +
  geom_hline(lty = 2, 
             alpha = 0.6,
             yintercept = 0) +
  scale_color_manual(name = '',
                     values = c('blue', 'orange'))

The above chart shows two lines: the orange line is based on Tukei's emperical results regarding the protection afforded by IRS against malaria infection (among children) as a function of time since IRS; the blue line is our simplification of Tukei's work.

@@

```{r, echo = TRUE}
weighter <- function(x){
   1/x
}
```

Where `x` is the distance (in kilometers) to the household whose "herd protection" score is being estimated. In other words, a household's "herd protection" score is the mean of all households' protection scores, weighted by the inverse of the distance away.

@@

## Density

We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels, weighted by the distance to the house in question. The below is an illustration of a house (the red x) and the neighboring houses (circles). The "weight" of each house is indicated by the circle size, and the protection level of the house is indicated by the shading of the circle.

```{r, fig.width = 8}
make_data <- function(n = 5){
  df <- data.frame(id = letters[1:n],
                 x = rnorm(n = n),
                 y = rnorm(n = n),
                 protection = round(sample(seq(0, 0.2, length = 100), n), digits = 1),
                 color = c('red', rep('grey', n-1))) %>%
  mutate(lng = x,
         lat = y)
  df$x[1] <- df$lng[1] <- df$lat[1] <- df$y[1] <- 0
library(sp)
coordinates(df) <- ~x+y
df$distance <- round(spDists(x = df)[1,], digits = 3) / 1
df$weight <- weighter(df$distance)
df$herd_protection <- df$protection * df$weight
return(df)
}

df <- make_data(n = 10)
make_plot <- function(df){
  ggplot(data = df@data,
       aes(x = lng,
           y = lat)) +
  geom_point(aes(size = weight,
                 color = protection)) +
  geom_point(data = df@data[1,],
             color = 'red',
             pch = 4,
             cex = 3) +
  theme_maragra() +
  labs(x = '',
       y = '') +
  theme(legend.position = 'right') +
  guides(color = guide_colourbar(title = 'Protection', title.position = 'top'),
         size = guide_legend(nrow = 5, title.position = 'top', nbin = 2)) +
  scale_color_continuous(low = 'orange', high = 'blue') +
    xlim(-2,2) +
    ylim(-2,2)
}
make_plot(df = df)

However, this consideration misses one important dimension: density. The below house's weighted average protection score is identical to that of the above house. In other words, at the time in question, at both locations, the percentage of houses with IRS coverage is the same, and the (weighted) average time since IRS is the same. However, the below house is likely much more protected than the above, given the _absolute_ number of nearby IRS-protected houses.

```{r, fig.width = 8}
df <- make_data(n = 100)
make_plot(df = df)
```

In the latter representation, the likelihood of a mosquito landing on a DDT-affected surface is much higher than the former, even though the relative IRS coverage for the neighborhood is identical. In order to account for this, we must take _density_ itself into account, ie the number of houses (and their protection level and proximity) within 1 kilometer.

Our approach is this: rather than using a weighted average of each household's individual protection score, we instead use a weighted sum. A particular household's herd protection score at any given time is the sum of all households' individual protection scores at that time multiplied by those households' distance weights.

## Combining them all together

The summation of (a) the IRS level of other houses at a certain time, (b) weighted by the distance of those houses to the house in question, (c) limited to only the 1 km radius multiplied yields a "herd protection score".

This score is conceptually similar to Cohen and Dupas' quantification of the positive externalities of ITN (bednet) use in Kenya [@Cohen2010] in that it attempts to estimate the protection conferred to "non-users" by users and uses a "weighted sum" (as opposed to average). Our approach differs in that the time dimension for IRS is much more important than ITN (ie, IRS is not a binary but rather a waning function, as described previously), requiring us to create herd scores for every day.

Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU]. No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects. Additionally, we using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007]

@@

### Table

@@

### HIV

We incorporated HIV prevalence of the area lived in, but it had not significance and was then removed from the model.

```{r}
plot(hiv_prevalence)
```

## Spatial protection surfaces

The below shows worker locations and the _average_ herd protection score for the entire study period.

```{r}
# Fortify bairros
library(maptools)
library(rgeos)
border <- gConvexHull(bairros_maragra_bairro)
border <- SpatialPolygonsDataFrame(Sr = border,
                                   data = data.frame(a = 1))
# border <- unionSpatialPolygons(bairros_maragra_bairro, 
#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
#                           threshold = 9^9)
border_fortified <- 
  broom::tidy(border,
              regions = 'Id')

# Create average herd risk score
x <- model_data %>%
  group_by(lng = longitude_aura,
           lat = latitude_aura) %>%
  summarise(herd = mean(herd, na.rm = TRUE)) %>%
  ungroup %>%
  filter(!is.na(lng))
x_sp <- x
coordinates(x_sp) <- ~lng+lat   
proj4string(x_sp) <- proj4string(border)
overs <- sp::over(x_sp, border)
x <- x[!is.na(overs),]
ggplot() +
  geom_point(data = x,
       aes(x = lng,
           y = lat,
           color = herd)) +
  geom_polygon(data = border_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = 'black') +
  ggthemes::theme_map() +
  scale_color_continuous(name = 'Herd\nprotection\nscore',
                         low = 'orange',
                         high = 'blue')

The below uses kernel density interpolation (via krieging) to estimate a generalized protection surface over the entire facility.

```{r}
# Create a gridded dataframe with values 
# for the entire range (bbox) of border
df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
                                 bbox(border)[1,2],
                                 by = 0.0008),
                       lat = seq(bbox(border)[2,1],
                                 bbox(border)[2,2],
                                 by = 0.0008),
                       herd = NA)
df_grid$latitude <- df_grid$lat
df_grid$longitude <- df_grid$lng
coordinates(df_grid) <- ~longitude+latitude

# Go through each row of df_grid, getting the 
# weighted mean irs score for that point
# and putting a color into df_grid
if('df_grid.RData' %in% dir()){
  load('df_grid.RData')
} else {
  for (i in 1:nrow(df_grid)){
  # Get distance from every point in x_sp
  distances <- spDistsN1(pts = x_sp,
                        pt = df_grid[i,],
                        longlat = TRUE)
  # Define which are acceptably close
  close_enough <- which(distances <= 50)
  # Get an herd score
  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
                       w = (1 / distances[close_enough]) ^2,
                       na.rm = TRUE)
  # Assign irs to the dataframe
  df_grid$herd[i] <- herd

}
save('df_grid', file = 'df_grid.RData')
}
# Convert df_grid to raster
temp <- df_grid@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
plot(r)

proj4string(df_grid) <- proj4string(border)
x <- over(df_grid, polygons(border))
df_grid_small <- df_grid[!is.na(x),]
temp <- df_grid_small@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
```

```{r}
plot(border)
plot(r, add = TRUE)
plot(border, add = TRUE)
```

### Theoretical protection surface

The below shows a theoretical protection surface assuming (a) uniform distribution of residences, (b) random distribution of IRS over time, but never going beyond 6 months between sprays.

```{r}
theoretical <- function(n){
  x <- lng <- runif(n = n, min = -1, max = 1)
y <- lat <- runif(n = n, min = -1, max = 1)
data <- data.frame(id = 1:n,
                   x,y, lng, lat,protection = sample(protection$protection[1:5], n,
                                                     replace = TRUE))
coordinates(data) <- ~x+y

# Loop through and get scores
out_list <- list()
counter <- 0
distances <- spDists(data)
for(j in 1:nrow(data)){
      sub_distances <- distances[j,]
      x <- data$protection
      w <- weighter(sub_distances)
      x <- x[is.finite(w)]

      # The weight
      w <- w[is.finite(w)]
      # Number of houses within 1k
      n <- length(which(sub_distances <= 1))
      # The weighted protection score (average)
      positivity <- stats::weighted.mean(x = x,
                                         w = w,
                                         na.rm = TRUE)
      # The weighted additive protection score
      s <- sum(x[sub_distances <= 1] * w[sub_distances <= 1], na.rm = TRUE)
      counter <- counter + 1

      # Update model_data
      out_data <- data_frame(id = data$id[j],
                             herd = positivity,
                             n = n,
                             s = s)
      out_list[[counter]] <- out_data
    }

x <- bind_rows(out_list)
x <- x %>%
    group_by(id) %>%
    summarise(herd = mean(herd, na.rm = TRUE),
              n = mean(n, na.rm = TRUE),
              s = mean(s, na.rm = TRUE))

data@data <- left_join(data@data, x)

border <- gConvexHull(data)
border <- SpatialPolygonsDataFrame(Sr = border,
                                   data = data.frame(a = 1))
border_fortified <- broom::tidy(border, regions = 1)

a1 <- ggplot() +
  geom_polygon(data = border_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               alpha = 0.2,
               color = 'black') +
  geom_point(data = data@data,
             aes(x = lng,
                 y = lat,
                 color = protection)) +
  theme_maragra() +
  scale_color_continuous(name = 'Individual\nprotection',
                         low = 'orange',
                         high = 'green',
                         guide = 'legend') +
  labs(x = '',
       y = '')
a2 <- ggplot() +
  geom_polygon(data = border_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               alpha = 0.2,
               color = 'black') +
  geom_point(data = data@data,
             aes(x = lng,
                 y = lat,
                 color = s)) +
  theme_maragra() +
  scale_color_continuous(name = 'Herd\nprotection',
                         low = 'orange',
                         high = 'green',
                         guide = 'legend') +
  labs(x = '',
       y = '')
Rmisc::multiplot(a1, a2, cols = 2)
}

#### 100 workers

```{r}
theoretical(n = 100) 
```

#### 1000 workers

```{r}
theoretical(n = 1000)
```

#### 3000 workers

```{r}
theoretical(n = 3000)
```

### 2018-04-28 inst/rmd/externalities.tex
big update

@@

To test whether IRS from one household has an effect on workers living
in another household, we first must make some assumptions about the
level and waning effect of IRS generally. For our purposes, we make a
rather simple assumption, based on the literature: that IRS has an
immediate effect on mosquitoes and should affect malaria risk following
the incubation period, that the waning effect begins in month 4 (Tukei
et al., 2017), that IRS effects those not living in the household by
reducing reproduction of mosquitoes {[}White2011{]}, and that IRS'
effect after 1 year is essentially null (Bukirwa et al., 2009). We
simplify these assumptions into a time-protection matrix, which can be
visualized in the below chart.

\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weighter <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    x[x }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\FloatTok{0.01}
\NormalTok{    out <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{/}\NormalTok{x)}\OperatorTok{^}\FloatTok{1.2}
    \KeywordTok{return}\NormalTok{(out)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Where \texttt{x} is the distance (in kilometers) to the household whose
``herd protection'' score is being estimated. In other words, a
household's ``herd protection'' score is the mean of all households'
protection scores, weighted by the inverse of the distance away raised
to the power of 1.2.

@@

\includegraphics{externalities_files/figure-latex/unnamed-chunk-5-1.pdf}

@@

\subsection{Incorporating into data}\label{incorporating-into-data}

We calculate a ``herd protection'' score for each location-date
combination.

\hypertarget{refs}{}
\hypertarget{ref-Bukirwa2009}{}
Bukirwa, H., Yau, V., Kigozi, R., Filler, S., Quick, L., Lugemwa, M.,
Dissanayake, G., Kamya, M., Wabwire-Mangen, F., Dorsey, G., 2009.
Assessing the impact of indoor residual spraying on malaria morbidity
using a sentinel site surveillance system in western uganda. American
Journal of Tropical Medicine and Hygiene 81, 611--614.
\url{https://doi.org/10.4269/ajtmh.2009.09-0126}

@@

>>>>

\usepackage{longtable}

@@

To test whether IRS from one household has an effect on workers living
in another household, we first must make some assumptions about the
level and waning effect of IRS generally. For our purposes, we make a
rather simple assumption, based on the literature: that IRS has an
immediate effect on mosquitoes and should affect malaria risk following
the incubation period, that the waning effect begins in month 4 (Tukei
et al., 2017), that IRS effects those not living in the household by
reducing reproduction of mosquitoes (White et al., 2011), and that IRS'
effect after 1 year is essentially null (Bukirwa et al., 2009). We
simplify these assumptions into a time-protection matrix, which can be
visualized in the below chart.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1} \end{center}

The above chart shows two lines: the orange line is based on Tukei's
emperical results regarding the protection afforded by IRS against
malaria infection (among children) as a function of time since IRS; the
blue line is our simplification of Tukei's work.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weighter <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x) \{}
    \DecValTok{1}\OperatorTok{/}\NormalTok{x}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Where \texttt{x} is the distance (in kilometers) to the household whose
``herd protection'' score is being estimated. In other words, a
household's ``herd protection'' score is the mean of all households'
protection scores, weighted by the inverse of the distance away.

@@

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-5-1} \end{center}

@@

\subsection{Density}\label{density}

We consider a house's ``herd'' protection level (ie, the protection
conferred to the house through externality) to be the average of the
other houses' non-herd protection levels, weighted by the distance to
the house in question. The below is an illustration of a house (the red
x) and the neighboring houses (circles). The ``weight'' of each house is
indicated by the circle size, and the protection level of the house is
indicated by the shading of the circle.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-6-1} \end{center}

However, this consideration misses one important dimension: density. The
below house's weighted average protection score is identical to that of
the above house. In other words, at the time in question, at both
locations, the percentage of houses with IRS coverage is the same, and
the (weighted) average time since IRS is the same. However, the below
house is likely much more protected than the above, given the
\emph{absolute} number of nearby IRS-protected houses.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-7-1} \end{center}

In the latter representation, the likelihood of a mosquito landing on a
DDT-affected surface is much higher than the former, even though the
relative IRS coverage for the neighborhood is identical. In order to
account for this, we must take \emph{density} itself into account, ie
the number of houses (and their protection level and proximity) within 1
kilometer.

Our approach is this: rather than using a weighted average of each
household's individual protection score, we instead use a weighted sum.
A particular household's herd protection score at any given time is the
sum of all households' individual protection scores at that time
multiplied by those households' distance weights.

\subsection{Combining them all
together}\label{combining-them-all-together}

The summation of (a) the IRS level of other houses at a certain time,
(b) weighted by the distance of those houses to the house in question,
(c) limited to only the 1 km radius multiplied yields a ``herd
protection score''.

This score is conceptually similar to Cohen and Dupas' quantification of
the positive externalities of ITN (bednet) use in Kenya (Cohen and
Dupas, 2010) in that it attempts to estimate the protection conferred to
``non-users'' by users and uses a ``weighted sum'' (as opposed to
average). Our approach differs in that the time dimension for IRS is
much more important than ITN (ie, IRS is not a binary but rather a
waning function, as described previously), requiring us to create herd
scores for every day.

Our approach can be justified in that previous studies have found strong
positive health externalities in malaria interventions related to ITN
coverage (Alaii et al., 2003). No studies exist on positive
externalities for IRS coverage, but to the extent that the mechanisms
for the reduction in infection are similar (reduction in the natural
reservoir of the disease, reduction in the number of vectors, etc.), it
is reasonable to assume similar effects. Additionally, we using weighted
distance for our score calculation (rather than simply defining a radius
threshold) because of previous studies' findings that there is a linear
decline in protection conferred to others with greater distance from an
intervention (Binka et al., 2007)

\subsection{Incorporating into data and
model}\label{incorporating-into-data-and-model}

We calculate a ``herd protection'' score for each location-date
combination (1759100 values). Whereas our original model specification
looked like:

\[
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
\]

Our new specification incorporates the herd protection value:

\[
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
\]

\(\hat{Y}_{it}\) is the rate of absence. \(\beta_{1}\) is the binary
``season'' variable, imputed from overall district clinical incidence.
Our intervention is whether the residence of the worker in question was
treated in the last year, and, if so, the time since treatment,
represented, respectively, by \(\beta_{2}\) and \(\beta_{3}\).
\(\beta_{4}\) \textbf{represents the distance-weighted herd protection
score}. \(\alpha_i\) represents the time invariant worker fixed effects,
and \(\delta_i\) represents the fixed effect of the particular malaria
season. \(\upsilon\) is the error term.

\subsection{Results}\label{results}

\subsubsection{Table}\label{table}

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-8}All absenteeism with herd immunity: model results}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}Malaria season & 3.597 (P<0.001)\\
\hspace{1em}Months since IRS: IRS & 6.149 (P<0.001)\\
\hspace{1em}Rainy day & 3.246 (P<0.001)\\
\hspace{1em}Herd protection & -0.01 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: IRS & -8.548 (P<0.001)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}Malaria season & 0.792 (P<0.001)\\
\hspace{1em}Months since IRS: IRS & 2.243 (P<0.001)\\
\hspace{1em}Rainy day & 4.552 (P<0.001)\\
\hspace{1em}Herd protection & -0.009 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: IRS & -1.508 (P=0.002)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}Malaria season & -0.362 (P<0.001)\\
\hspace{1em}Months since IRS: IRS & -0.229 (P=0.004)\\
\hspace{1em}Rainy day & 0.037 (P=0.261)\\
\hspace{1em}Herd protection & 0.002 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: IRS & 0.26 (P=0.014)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}Malaria season & 3.151 (P<0.001)\\
\hspace{1em}Months since IRS: IRS & -1.718 (P=0.076)\\
\hspace{1em}Rainy day & 0.224 (P=0.469)\\
\hspace{1em}Herd protection & 0.002 (P=0.772)\\
\hspace{1em}Malaria season:Months since IRS: IRS & 2.657 (P=0.02)\\*
\end{longtable}

\subsubsection{HIV}\label{hiv}

We incorporated HIV prevalence of the area lived in, but it had not
significance and was then removed from the model.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-9-1} \end{center}

\subsection{Spatial protection
surfaces}\label{spatial-protection-surfaces}

The below shows worker locations and the \emph{average} herd protection
score for the entire study period.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-10-1} \end{center}

The below uses kernel density interpolation (via krieging) to estimate a
generalized protection surface over the entire facility.

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-11-1} \end{center}

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-12-1} \end{center}

\subsubsection{Theoretical protection
surface}\label{theoretical-protection-surface}

The below shows a theoretical protection surface assuming (a) uniform
distribution of residences, (b) random distribution of IRS over time,
but never going beyond 6 months between sprays.

\paragraph{100 workers}\label{workers}

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-14-1} \end{center}

\paragraph{1000 workers}\label{workers-1}

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-15-1} \end{center}

\paragraph{3000 workers}\label{workers-2}

\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-16-1} \end{center}

\subsection{Conclusion}\label{conclusion}

Using the approach laid out here, there appears to be a significant
protective effect afforded by the IRS of neighbors. If this approach is
deemed correct, then next steps are determining:

\begin{itemize}
\tightlist
\item
  How to translate the effect into costs and savings (simulation based
  approach at various IRS coverage levels?).
\item
  How to combine this effect with the other potential spillover effect
  (ie, core-periphery issues).
\item
  How / where to report on this in the paper (is this a ``method'' or
  ``result''?).
\end{itemize}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Alaii2003FactorsAU}{}
Alaii, J.A., Hawley, W.A., Kolczak, M.S., Kuile, F.O.T., Gimnig, J.E.,
Vulule, J., Odhacha, A., Oloo, A.J., Nahlen, B.L., Phillips-Howard,
P.A., 2003. Factors affecting use of permethrin-treated bed nets during
a randomized controlled trial in western kenya. The American journal of
tropical medicine and hygiene 68 4 Suppl, 137--41.

\hypertarget{ref-Binka2007}{}
Binka, F.N., Kubaje, A., Adjuik, M., Williams, L.A., Lengeler, C.,
Maude, G.H., Armah, G.E., Kajihara, B., Adiamah, J.H., Smith, P.G.,
2007. Impact of permethrin impregnated bednets on child mortality in
kassena-nankana district, ghana: A randomized controlled trial. Tropical
Medicine \& International Health 1, 147--154.
\url{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}

\hypertarget{ref-Bukirwa2009}{}
Bukirwa, H., Yau, V., Kigozi, R., Filler, S., Quick, L., Lugemwa, M.,
Dissanayake, G., Kamya, M., Wabwire-Mangen, F., Dorsey, G., 2009.
Assessing the impact of indoor residual spraying on malaria morbidity
using a sentinel site surveillance system in western uganda. American
Journal of Tropical Medicine and Hygiene 81, 611--614.
\url{https://doi.org/10.4269/ajtmh.2009.09-0126}

\hypertarget{ref-Cohen2010}{}
Cohen, J., Dupas, P., 2010. Free distribution or cost-sharing? Evidence
from a randomized malaria prevention experiment. Quarterly Journal of
Economics 125, 1--45. \url{https://doi.org/10.1162/qjec.2010.125.1.1}

@@

\hypertarget{ref-White2011}{}
White, M.T., Griffin, J.T., Churcher, T.S., Ferguson, N.M., Basáñez,
M.-G., Ghani, A.C., 2011. Modelling the impact of vector control
interventions on anopheles gambiae population dynamics. Parasites \&
Vectors 4, 153. \url{https://doi.org/10.1186/1756-3305-4-153}
